Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-23 RNARedPrint.tex
Meeting Seb

\subsection*{Effectively targeting Turner energies}
We used our Boltzmann sampling strategy (Algorithms
\ref{alg:pf} and \ref{alg:sampling}), to sample valid sequences for
given target structures and weights $\pi_1,\dots,\pi_k$.  Moreover, we
used multi-dimensional Boltzmann sampling to target specific energies and
\GCb\%.  Our tool \ourprog{} evaluates energies according to the
stacking energy model $\EnergyStacking$, whose parameters were fitted
to best approximate Turner energies. As well, we implemented and
fitted a base pair energy model for \ourprog{}, which was not studied
for its targeting performance (both models:
Supp. Mat.~\ref{appsec:modelparameters}).
%
%\paragraph{Implementation of multi-dimensional Boltzmann sampling}
%As suggested in Section~\ref{sec:multiBoltzmann}, we heuristically optimize weights to target specific energies and \GCb\% in an iterative procedure (based on the core sampling algorithm).
%Specifically, energy weights are initialized as $w_\ell=e^{\frac{1}{RT}}$; the gc weight as $w_{gc}=1$. For each iteration, $\psi\cdot n$ sequences are sampled with the current weights. Then, we determine the mean energies $E(S;R_\ell)$ and the mean GC content over the sampled sequences. For the next iteration, the weights are updated by the rules $w_\ell = w_\ell \gamma^{mean(E(S;R_\ell))-E_\ell}$ and $w_{gc} = w_{gc}\frac{GC_{target}}{mean(GC)}$. During this procedure, we collect   in a set $P$ all the sequences exhibiting energies within $[E_1-\delta,E_1+\delta]\times [E_2-\delta,E_2+\delta]\ldots$ and and GC-content within $[GC_{target}-\idelta',GC_{target}+\delta']$.
%If $|P| > n$ or an maximum amount of iterations is reached, the sequences in $P$ are returned.
%

Fig.~\ref{fig:energydist-pk} illustrates how well complex realistic energy models can be approximated based on simpler, but better tractable ones. For
the two target structures of Fig.~\ref{fig:energydist-pk}B,
Fig.~\ref{fig:energydist-pk}A shows the good fit between realistic
energies in the full-fledged Dirks and Pierce energy model for
pseudoknots (D\&P model) and energies in the stacking energy
model. For the shown fits we sampled $n=1\,000$ sequences, targeting a
\GCb\% of $60\%$.
%
For an example instance of the \texttt{Modena} benchmark with two
pseudoknotted target structures, Fig.~\ref{fig:energydist-pk}B shows
the Turner energy distributions of the single structures as they
result from sampling with different weight parameters. The figure
illustrates how our multidimensionnal Boltzmann sampling strategy can,
to a large extent, independently shift the Turner energies of sampled
sequences towards prescribed targets. See Supp. Mat.,
Fig.~\ref{fig:energydist} for a further example with three
pseudoknot-free target structures.

>>>>

\subsection*{A parsimonious stacking energy model as a proxy for the Turner energy model}
Sampling for fixed weight
Time benchmark

Figure 10 - Correlation plot -- Right
Figure 2 -- Distribution plot -- Right (Turner is even untractable to compute the tree width itself in benchmark instances)
Figure 3 -- Right -- Distribution

\subsection*{Effectively targeting Turner energies using multidimensionnal sampling}
\revised{

We used our Boltzmann sampling strategy (Algorithms
\ref{alg:pf} and \ref{alg:sampling}), to sample valid sequences for
given target structures and weights $\pi_1,\dots,\pi_k$.  Moreover, we
used multi-dimensional Boltzmann sampling to target specific energies and
\GCb\%.  Our tool \ourprog{} evaluates energies according to the
stacking energy model $\EnergyStacking$, whose parameters were fitted
to best approximate Turner energies. As well, we implemented and
fitted a base pair energy model for \ourprog{}, which was not studied
for its targeting performance (both models:
Supp. Mat.~\ref{appsec:modelparameters}).
%
%\paragraph{Implementation of multi-dimensional Boltzmann sampling}
%As suggested in Section~\ref{sec:multiBoltzmann}, we heuristically optimize weights to target specific energies and \GCb\% in an iterative procedure (based on the core sampling algorithm).
%Specifically, energy weights are initialized as $w_\ell=e^{\frac{1}{RT}}$; the gc weight as $w_{gc}=1$. For each iteration, $\psi\cdot n$ sequences are sampled with the current weights. Then, we determine the mean energies $E(S;R_\ell)$ and the mean GC content over the sampled sequences. For the next iteration, the weights are updated by the rules $w_\ell = w_\ell \gamma^{mean(E(S;R_\ell))-E_\ell}$ and $w_{gc} = w_{gc}\frac{GC_{target}}{mean(GC)}$. During this procedure, we collect   in a set $P$ all the sequences exhibiting energies within $[E_1-\delta,E_1+\delta]\times [E_2-\delta,E_2+\delta]\ldots$ and and GC-content within $[GC_{target}-\idelta',GC_{target}+\delta']$.
%If $|P| > n$ or an maximum amount of iterations is reached, the sequences in $P$ are returned.
%

Fig.~\ref{fig:energydist-pk} illustrates how well complex realistic energy models can be approximated based on simpler, but better tractable ones. For
the two target structures of Fig.~\ref{fig:energydist-pk}B,
Fig.~\ref{fig:energydist-pk}A shows the good fit between realistic
energies in the full-fledged Dirks and Pierce energy model for
pseudoknots (D\&P model) and energies in the stacking energy
model. For the shown fits we sampled $n=1\,000$ sequences, targeting a
\GCb\% of $60\%$.
%
For an example instance of the \texttt{Modena} benchmark with two
pseudoknotted target structures, Fig.~\ref{fig:energydist-pk}B shows
the Turner energy distributions of the single structures as they
result from sampling with different weight parameters. The figure
illustrates how our multidimensional Boltzmann sampling strategy can,
to a large extent, independently shift the Turner energies of sampled
sequences towards prescribed targets. See Supp. Mat.,
Fig.~\ref{fig:energydist} for a further example with three
pseudoknot-free target structures.}
